maxent model avec xvalidation (year-based test training)

### 2019-04-18 src/paper2_maxent_avec_xvalidation.Rmd
paper 2 runs work.

y1998 = 2.75
y1999 = 2.25
y2000 = 2
y2001 = 1.75
y2002 = 2
y2003 = 2
y2004 = 2
y2005 = 2.25
y2006 = 1.75
y2007 = 1.75
y2008 = 2.25
y2009 = 2.25
y2010 = 2.5
y2011 =  2.5

```{r final models plus predict}
modelmonth <- cbind(paste0(mthna, unique(prback$year)))
aicetc <- read.csv("../data/env/AICcFinalmod.csv", header = TRUE) #pre-created csv file
unlayer <- read.csv(paste0("../data/env/asc_layerlist", mthna, ".csv"), header = TRUE) #if code doesnt work put back to unlayer_yr

for (year in 1:nrow(modelmonth)){
  timeslice <- modelmonth[year] #testing timeslice
  yr <- str_sub(timeslice, -4) #the year being used for testing
  train <- subset(prback, year != yr) #training data is all years but testing
  trainpa <- train$occurrence #creates a vector of presence/background ID (1/0)
  test <-  subset(prback, year == yr) #creates the testing dataset
  testp <- subset(test, occurrence == "1") #splits the testing dataset into presences
  testp <- subset(testp, select = -c(year, occurrence)) #don't need year and occurrence for modelling
  testb <- subset(test, occurrence == "0") #splits the testing dataset into background
  testb <- subset(testb, select = -c(year, occurrence)) #don't need year and occurrence for modelling
  train  <- subset(train, select = -c(year, occurrence)) #don't need year and occurrence for modelling

  if (timeslice == "y2001" | timeslice == "y2006" | timeslice == "y2007"){
    betaval <- 1.75
  } else if (timeslice == "y2000"| timeslice == "y2002" | timeslice == "y2003" | timeslice == "y2004"){
    betaval <- 2
  } else if (timeslice == "y1999" | timeslice == "y2005" | timeslice == "y2008" | timeslice == "y2009"){
    betaval <- 2.25
  } else if (timeslice == "y2010" | timeslice == "y2011"){
    betaval <- 2.5
  } else if (timeslice == "y1998"){
    betaval <- 2.75
  }

    mod <- maxent(x = train, p = trainpa, path = paste0("../output/maxent/p2/models/", timeslice), args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", paste0("betamultiplier=", betaval)))

@@

    #count number of non-zero coefficients, excluding the final four rows
    coefile <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/species.lambdas"), header = FALSE, stringsAsFactors=FALSE)
    coefile <- head(coefile, -4)
    coefile2 <- subset(coefile, V2 > 0 | V2 < 0)
    nocoeff <- nrow(coefile2)
    bick <- sum(as.numeric(coefile2[, 2]))
    bic <- (bick*log(sum(trainpa == 1)) - 2*11)

    #no presences used in model (less those used for testing)
    maxresults <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    notrainingpoints <- maxresults$X.Training.samples
    testauc <- maxresults$Test.AUC
    trainauc <- maxresults$Training.AUC

    #natural log
    samplepred <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/species_samplePredictions.csv"), header = TRUE)
    samplepred$nat_log <- log(samplepred$Raw.prediction)
    sumsamplepred <- sum(samplepred$nat_log)

    #test auc scores and other lovely things you might want to include
    ev <- evaluate(p = testp, a= testb, model = mod)
    evauc <- ev@auc
    evcor <- ev@cor
    evpcor <- ev@pcor
    evnp <- ev@np
    evna <- ev@na
    evmaxtprtnr <- threshold(ev, 'spec_sens')
    evmaxkap <- threshold(ev, 'kappa')
    evno_omission <- threshold(ev, 'no_omission')
    evprevalence <- threshold(ev, 'prevalence')
    evequal_sens_spec <- threshold(ev, 'equal_sens_spec')
    evsensitivity <- threshold(ev, 'sensitivity')

      aicetc$train_AUC <- trainauc
      aicetc$no_trainng_points <- notrainingpoints
      aicetc$test_coefficients <- nocoeff
      aicetc$test_sum.of.logs <- sumsamplepred
      aicetc$test_AUC <- evauc
      aicetc$test_evcor <- evcor
      aicetc$test_evpcor <- evpcor
      aicetc$test_nopresence <- evnp
      aicetc$test_nobackground <- evna
      aicetc$test_maxsumsenspec <- evmaxtprtnr
      aicetc$test_maxkap <- evmaxkap
      aicetc$test_no_omission <- evno_omission
      aicetc$test_prevalence <- evprevalence
      aicetc$test_equal_sens_spec <- evequal_sens_spec
      aicetc$test_sensitivity <- evsensitivity
      aicetc$test_bic <- bic #not sure this is correct...
      aicetc$beta <- betaval

@@

    write.csv(aicetc, paste0("../output/maxent/p2/models/", timeslice, "/AICforMaxentReg.csv"), row.names = FALSE)

@@

    for (td in 1:nrow(unlayer_yr)){
      temp_depth <- raster(unlayer_yr$temp_depth[td])
      crs(temp_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(temp_depth) <- "temp_depth"
      temp_depth <- projectRaster(temp_depth, aea)
      chl_surface <- raster(unlayer_yr$chl_surface[td])
      crs(chl_surface) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(chl_surface) <- "chl_surface"
      chl_surface <- projectRaster(chl_surface, aea)
      o2_depth <- raster(unlayer_yr$o2_depth[td])
      crs(o2_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(o2_depth) <- "o2_depth"
      o2_depth <- projectRaster(o2_depth, aea)
      salinity_depth <- raster(unlayer_yr$salinity_depth[td])
      crs(salinity_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(salinity_depth) <- "salinity_depth"
      salinity_depth <- projectRaster(salinity_depth, aea)
      nao_sample <- raster(unlayer_yr$nao_sample[td])
      crs(nao_sample) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(nao_sample) <- "nao_sample"
      nao_sample <- projectRaster(nao_sample, aea)
      nao_prev <- raster(unlayer_yr$nao_prev[td])
      crs(nao_prev) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(nao_prev) <- "nao_prev"
      nao_prev <- projectRaster(nao_prev, aea)
      nao_winter <- raster(unlayer_yr$nao_winter[td])
      crs(nao_winter) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
      names(nao_winter) <- "nao_winter"
      nao_winter <- projectRaster(nao_winter, aea)
      amo_sample <- raster(unlayer_yr$amo_sample[td])
      crs(amo_sample) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(amo_sample) <- "amo_sample"
      amo_sample <- projectRaster(amo_sample, aea)
      amo_winter <- raster(unlayer_yr$amo_winter[td])
      crs(amo_winter) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
      names(amo_winter) <- "amo_winter"
      amo_winter <- projectRaster(amo_winter, aea)
      stackn <- unlayer_yr$stackname[td]
      stack <- stack(temp_depth, salinity_depth, o2_depth, chl_surface, nao_sample, nao_prev, nao_winter, amo_sample, amo_winter)
      #stackSave(test,unlayer_yr$stackname[td])
      pred <- predict(mod, stack, file = paste0("../output/maxent/p2/predictions/", mthna, "/", "cloglog_", stackn, ".asc"), overwrite=TRUE, args='outputformat=cloglog') 
      predcum <- predict(mod, stack, file = paste0("../output/maxent/p2/predictions/", mthna, "/", "cumu_", stackn, ".asc"), overwrite=TRUE, args=c('outputformat=Cumulative')) 
      prplot <- plot(pred, png(paste0("../output/maxent/p2/predictions/", mthna, "/", "cloglog_", stackn, ".png")))
      dev.off() # stops automatic saving of the plot to a png
      prplotcum <- plot(predcum, png(paste0("../output/maxent/p2/predictions/", mthna, "/", "cumu_", stackn, ".png"))) 
      dev.off() # stops automatic saving of the plot to a png
    }
}

@@

- the AICforMaxentReg.csv 
- the maxentResults.csv
- the asc files per depth layer

open all maxentReg files

```{r avg maxentReg}

for (i in 1:length(modelmonth)){
  timeslice <- modelmonth[i]
  yr <- str_sub(timeslice, -4) #the year being used for testing
  if(yr == 1998){
    aic1998 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr1998 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 1999){
    aic1999 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr1999 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2000){
    aic2000 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2000 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2001){
    aic2001 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2001 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2002){
    aic2002 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2002 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2003){
    aic2003 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2003 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2004){
    aic2004 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2004 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2005){
    aic2005 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2005 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2006){
    aic2006 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2006 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2007){
    aic2007 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2007 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2008){
    aic2008 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2008 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2009){
    aic2009 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2009 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2010){
    aic2010 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2010 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2011){
    aic2011 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2011 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2012){
    aic2012 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2012 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2013){
    aic2013 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2013 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2014){
    aic2014 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2014 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)
    } else if (yr == 2015){
    aic2015 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/AICforMaxentReg.csv"), header = TRUE)
    maxr2015 <- read.csv(paste0("../output/maxent/p2/models/", timeslice, "/maxentResults.csv"), header = TRUE)

aic <- rbind(aic1998, aic1999, aic2000, aic2001, aic2002, aic2003, aic2004, aic2005, aic2006, aic2007, aic2008, aic2009, aic2010, aic2011)
maxr <- rbind(maxr1998, maxr1999, maxr2000, maxr2001, maxr2002, maxr2003, maxr2004, maxr2005, maxr2006, maxr2007, maxr2008, maxr2009, maxr2010, maxr2011)

aicm <- colMeans(aic)
aicm$mod <- "avg"
avgAIC <- rbind(aic, aicm)

maxr$Species <- 1
maxrm <- colMeans(maxr)
maxrm$Species <- "avg"
maxrm <- rbind(maxr, maxrm)

write.csv(avgAIC, paste0("../output/maxent/p2/models/", mthna, "avgAIC.csv"), row.names = FALSE)
write.csv(maxrm, paste0("../output/maxent/p2/models/", mthna, "avgmaxentresults.csv"), row.names = FALSE)
```

@@

>>>>

2003: 0.25
2004: 1.5
2005: 0.75
2006: 2.25

```{r final models plus predict}
modelmonth <- cbind(paste0(mthna, unique(prback$year)))
aicetc <- read.csv("../data/env/AICcCalculations_p2.csv", header = TRUE) #pre-created csv file
unlayer <- read.csv(paste0("../data/env/asc_layerlist", mthna, ".csv"), header = TRUE) #if code doesnt work put back to unlayer_yr
nfolds <- cbind(c(1:5))

for (year in 1:nrow(modelmonth)){
  timeslice <- modelmonth[year] #testing timeslice
  yr <- str_sub(timeslice, -4) #the year being used for testing
  train <- subset(prback, year == yr) #training data is all years but testing
  trainpa <- train$occurrence #creates a vector of presence/background ID (1/0)
  train  <- subset(train, select = -c(year, occurrence)) #don't need year and occurrence for modelling

  if (yr == "2003"){
    betaval <- 0.25
  } else if (yr == "2004"){
    betaval <- 1.5
  } else if (yr == "2005"){
    betaval <- 0.75
  } else if (yr == "2006"){
    betaval <- 2.25
  }

  mod <- maxent(x = train, p = trainpa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=8", "addsamplestobackground=TRUE", paste0("replicates=", nrow(nfolds)), "replicatetype=crossvalidate", paste0("betamultiplier=", betaval)), path = paste0("C:/Users/bunny/Documents/GitHub/Chapter_1/output/maxent/p2/models/", timeslice)) #here the full path is needed

@@

      aicetc$bic <- bic #not sure this is correct...
      aicetc$beta <- betaval
      aicetc$foldno <- modn

@@

    aicetc$AIC <- ((2*aicetc$coefficients)-(2*aicetc$sum.of.logs))
    aicetc$AICc <- (-2*(aicetc$sum.of.logs)+((2*aicetc$coefficients)*(aicetc$n/(aicetc$n - aicetc$coefficients-1))))
    write.csv(aicetc, paste0("../output/maxent/p2/models/", timeslice, "/", betaval, "_", modn, "_", "AICforMaxentReg.csv"), row.names = FALSE)
      }

    aicperbetalst <- list.files(paste0("../output/maxent/p2/models/", timeslice, "/"), pattern = paste0("^", betaval, ".*AICforMaxentReg.csv"), full.names = TRUE)
    aicperbeta <- lapply(aicperbetalst, read.csv, header = TRUE)
    aicperbeta <- do.call(rbind, aicperbeta)
    aicbetamean <- colMeans(aicperbeta)
    aicperbeta <- rbind(aicperbeta, aicbetamean)
    aicperbeta$foldno[f+1] <- f+1
    write.csv(aicperbeta, paste0("../output/maxent/p2/models/", timeslice, "/final_model_aicperbetasum_", betaval, ".csv"), row.names = FALSE)

@@

    for (td in 1:nrow(unlayer_yr)){
      temp_depth <- raster(unlayer_yr$temp_depth[td])
      crs(temp_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(temp_depth) <- "temp_depth"
      temp_depth <- projectRaster(temp_depth, aea)
      chl_surface <- raster(unlayer_yr$chl_surface[td])
      crs(chl_surface) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(chl_surface) <- "chl_surface"
      chl_surface <- projectRaster(chl_surface, aea)
      o2_depth <- raster(unlayer_yr$o2_depth[td])
      crs(o2_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(o2_depth) <- "o2_depth"
      o2_depth <- projectRaster(o2_depth, aea)
      salinity_depth <- raster(unlayer_yr$salinity_depth[td])
      crs(salinity_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(salinity_depth) <- "salinity_depth"
      salinity_depth <- projectRaster(salinity_depth, aea)
      nao_sample <- raster(unlayer_yr$nao_sample[td])
      crs(nao_sample) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(nao_sample) <- "nao_sample"
      nao_sample <- projectRaster(nao_sample, aea)
      nao_prev <- raster(unlayer_yr$nao_prev[td])
      crs(nao_prev) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(nao_prev) <- "nao_prev"
      nao_prev <- projectRaster(nao_prev, aea)
      nao_winter <- raster(unlayer_yr$nao_winter[td])
      crs(nao_winter) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
      names(nao_winter) <- "nao_winter"
      nao_winter <- projectRaster(nao_winter, aea)
      amo_sample <- raster(unlayer_yr$amo_sample[td])
      crs(amo_sample) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
      names(amo_sample) <- "amo_sample"
      amo_sample <- projectRaster(amo_sample, aea)
      amo_winter <- raster(unlayer_yr$amo_winter[td])
      crs(amo_winter) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
      names(amo_winter) <- "amo_winter"
      amo_winter <- projectRaster(amo_winter, aea)
      stackn <- unlayer_yr$stackname[td]
      stack <- stack(temp_depth, salinity_depth, o2_depth, chl_surface, nao_sample, nao_prev, nao_winter, amo_sample, amo_winter)
      #stackSave(test,unlayer_yr$stackname[td])
      pred <- predict(mod, stack, file = paste0("../output/maxent/p2/predictions/", timeslice, "/", "cloglog_", stackn, ".asc"), overwrite=TRUE, args='outputformat=cloglog') 
      predcum <- predict(mod, stack, file = paste0("../output/maxent/p2/predictions/", timeslice, "/", "cumu_", stackn, ".asc"), overwrite=TRUE, args=c('outputformat=Cumulative')) 
      for (pf in 1:nrow(nfolds)){
        plotn <- nfolds[pf]-1
        lyr <- nfolds[pf]
        prplot <- plot(pred[[pf]], png(paste0("../output/maxent/p2/predictions/", timeslice, "/", "cloglog_", stackn, "_", plotn, ".png")))
        dev.off() # stops automatic saving of the plot to a png
      }
    }
}

@@

- the asc files per depth layer per year

@@

NEEDS TO BE DONE FROM HERE!!
